Reconstructing Propositional Proofs in Type Theory

### 2017-10-27 paper/sections/metis-language-and-proofs.tex
[ paper ] minor fixes.

The \TPTP language is the input language to encode problems used by \Metis.
It includes the first-order form (denoted by \name{fof}) and clause normal form (denoted by \name{cnf}) formats~\cite{sutcliffe2009}.
The \TPTP syntax describes a well-defined grammar to handle annotated
formulas with the following form:
\begin{verbatim}
  language(name, role, formula).,
\end{verbatim}
where the \name{language} can be \name{fof} or \name{cnf}. The
\name{name} serves to identify the formula within the problem. Each
formula assumes a \name{role}, this could be an \name{axiom},
\name{conjecture}, \name{definition}, \name{plain} or an
\name{hypothesis}.  The formulas include the constants
\verb!$true! and \verb!$false!, the negation unary operator~\verb!~!,
and the binary connectives \verb!&!, \verb!|!, \verb!=>!, and
\verb!<=>! to represent $⊤$, $⊥$, $¬$, $∧$, $∨$, $⇒$, and $⇔$
respectively.

\begin{myexamplenum}
  For instance, let us express the problem
  $p\, ⊢ ¬ (p ∧ ¬ p) ∨ (q ∧ ¬ q)$ in \TPTP syntax. We begin by
  declaring the $p$ axiom using the \verb!axiom!  keyword. Next, we
  include the expected conclusion using the \verb!conjecture! keyword.
\begin{verbatim}
  fof(h, axiom, p).
  fof(goal, conjecture, ~ ((p & ~ p) | (q & ~ q))).
\end{verbatim}
\end{myexamplenum}

@@

\begin{verbatim}
  language(name, role, formula, source [,useful info]).
\end{verbatim}

@@

We adopt a customized \TSTP syntax to keep as short as possible the \Metis
derivations for increasing the readability in this paper.

>>>>

The \TPTP language is the input language to encode problems used by \Metis.
It includes the first-order form (denoted by \name{fof}) and clause normal form (denoted by \name{cnf}) formats~\cite{sutcliffe2009}.
The \TPTP syntax describes a well-defined grammar to handle annotated
formulas with the following form:

\begin{verbatim}
  language(name, role, formula).,
\end{verbatim}

where the \name{language} can be \name{fof} or \name{cnf}. The
\name{name} serves to identify the formula within the problem. Each
formula assumes a \name{role}, this could be an \name{axiom},
\name{conjecture}, \name{definition}, \name{plain} or an
\name{hypothesis}.  The formulas include the constants
\verb!$true! and \verb!$false!, the negation unary operator~(\verb!~!),
and the binary connectives \verb!&!, \verb!|!, \verb!=>!, and
(\verb!<=>!) to represent ($⊤$, $⊥$, $¬$, $∧$, $∨$, $⇒$), and ($⇔$)
respectively.

\begin{myexamplenum}
  For instance, let us express the problem
  $p\, ⊢ ¬ (p ∧ ¬ p) ∨ (q ∧ ¬ q)$ in \TPTP syntax. We begin by
  declaring the $p$ axiom using the \verb!axiom!  keyword. Next, we
  include the expected conclusion using the \verb!conjecture! keyword.

\begin{verbatim}
  fof(h, axiom, p).
  fof(goal, conjecture, ~ ((p & ~ p) | (q & ~ q))).
\end{verbatim}

\end{myexamplenum}

@@

\begin{verbatim}
  language(name, role, formula, source [,useful info]).,
\end{verbatim}

@@

We adopt a customized \TSTP syntax to keep as short as possible the \Metis
derivations for increasing the readability of this paper.

### 2017-10-27 paper/sections/type-theory.tex
[ paper ] removed CPL.
[ paper ] minor fixes.

To reconstruct proofs generated by \Metis, we have formalized in type
theory the classical propositional logic.  A briefly description of it
is presented in Appendix~\ref{app:CPL}. In this formalization, we
assume the principle of excluded third (henceforth \abbre{PEM}) as the
only axiom since \Metis is a prover for classical logic. Assuming
\abbre{PEM}, we can justify refutation proofs by deriving from it the
\emph{reductio ad absurdum} rule (henceforth \abbre{RAA}). The
\abbrev{RAA} rule is the formulation of the principle of proof by
contradiction, that is, a derivation of a contradiction,~$⊥$, from the
hypothesis~$¬ φ$, is a derivation of~$φ$.

@@

We implemented in~\cite{AgdaProp} the syntactical consequence relation in a
similar way as it was presented in~\cite{Altenkirch2015}. For that reason, we
have included structural rules like \emph{weaken}, formation and elimination
rules for connectives and the PEM axiom as the constructors of the theorem type
in Fig.~\ref{fig:CPL-inference-rules}.

>>>>

To reconstruct proofs generated by \Metis, we have formalized in type theory the
classical propositional logic in~\cite{AgdaProp}.  A briefly description of it
is presented in \cite{VanDalen1994}. In this formalization, we assume the
principle of excluded third (henceforth \abbre{PEM}) as the only axiom since
\Metis is a prover for classical logic. Assuming \abbre{PEM}, we can justify
refutation proofs by deriving from it the \emph{reductio ad absurdum} rule
(henceforth \abbre{RAA}). The \abbrev{RAA} rule is the formulation of the
principle of proof by contradiction, that is, a derivation of a
contradiction,~$⊥$, from the hypothesis~$¬ φ$, is a derivation of~$φ$.

@@

We implemented in~\cite{AgdaProp} the syntactical consequence relation in a
similar way as it was presented in~\cite{Altenkirch2015}. For that reason, we
have included structural rules like \emph{weaken}, formation and elimination
rules for connectives and the PEM axiom as the valid inference rules
in Fig.~\ref{fig:CPL-inference-rules}.
